---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Nepal-Conflict-and-Elections

<!-- badges: start -->

<!-- badges: end -->


## Introduction

Nepal is one of only a few countries in the world that successfully integrated a rebel group into its political system, a process that has yielded relatively peaceful results to the present day. From 1996 to 2006, the Nepalese people faced the tragedy of a [civil war](https://en.wikipedia.org/wiki/Nepalese_Civil_War) that destroyed lives, communities and families alike. For the personal stories of those affected, the [INSEC](https://www.insec.org.np/victim) database comprehensively catalogues details on the thousands of dead, wounded and missing of the conflict.



Given Nepal's unique story, I put together a study to examine the relationship between a rebel group's actions during a civil conflict (regarding their treatment of civilians and casualties inflicted on government forces), and its performance in postwar transitional elections. 


## The data

Here is a subsection of the dataset I compiled. There is also data on Maoist electoral performance, and some attributes of the successful and unsuccessful candidates who contested the single-member constituencies. 

```{r}
library(tidyverse)
library(kableExtra)

df = read.csv("completedata.csv")
kable(head(df))
```



## Descriptive statistics


With this data, we can examine the mean age of candidates and winning candidates in each electoral district along with concurrent violence statistics.

```{r}
mean(df$winnerage)
mean(df$candidate_avgage)
mean(df$civdeath_gov)
mean(df$civdeath_mao)
mean(df$govtdeath)
mean(df$maodeath)
mean(df$statebased)
mean(df$onesided)




```


## Plots

We can visualize how the age of winning candidates moves with gender and the Maoist vote share in each electoral district.

```{r}
ggplot(df, aes(maovote, winnerage, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```


This plot regresses turnout with the average age of the winning candidate separated by gender. We can see that there is a much stronger positive relationship between turnout and winner age in constituencies with female winning candidates as a pose to male candidates.

```{r}
ggplot(df, aes(turnout, winnerage, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```


This plot regresses Maoist performance with the average age of the candidates contesting single-member constituencies disaggregated by gender.

```{r}
ggplot(df, aes(maovote, candidate_avgage, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

Now, moving onto regressions that involve conflict statistics.

```{r}
ggplot(df, aes(maodeath, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```