---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Nepal-Conflict-and-Elections

<!-- badges: start -->

<!-- badges: end -->


## Introduction

Nepal is one of only a few countries in the world that successfully integrated a rebel group into its political system, a process that has yielded relatively peaceful results to the present day. From 1996 to 2006, the Nepalese people faced the tragedy of a [civil war](https://en.wikipedia.org/wiki/Nepalese_Civil_War) that destroyed lives, communities and families alike. For the personal stories of those affected, the [INSEC](https://www.insec.org.np/victim) database comprehensively catalogues details on the thousands of dead, wounded and missing of the conflict.



Given Nepal's unique story, I put together a study to examine the relationship between a rebel group's actions during a civil conflict (regarding their treatment of civilians and casualties inflicted on government forces), and its performance in postwar transitional elections. 


## The data

Here is a subsection of the dataset I compiled. There is also data on Maoist electoral performance, and some attributes of the successful and unsuccessful candidates who contested the single-member constituencies. 

```{r}
library(tidyverse)
library(kableExtra)
library(psych)

df = read.csv("completedata.csv")
kable(head(df))
```



## Descriptive statistics


With this data, we can examine the mean age of winning candidates in each electoral district along with the mean level of civilian killed by Maoists per 100,000 eligible voters and the mean level of Maoist rebels killed by government forces per 100,000 eligible voters.

```{r}

df %>% summarise(civdeath_mao = mean(df$civdeath_mao), maodeath = mean(df$maodeath), winnerage = mean(df$winnerage))%>% tibble()

```

Key

civdeath_gov -> Civilian deaths perpetrated by Government Forces (Royal Nepalese Army, Nepal Police, Armed Police Force) per 100,000 eligible voters

civdeath_mao -> Civilian deaths perpetrated by the Communist Party of Nepal (Maoist) per 100,000 eligible voters

govtdeath -> Government deaths perpetrated by the Maoist rebels per 100,000 eligible voters

maodeath -> Maoist deaths perpetrated by Government forces per 100,000 eligible voters.

statebased -> State-based deaths (govtdeath + maodeath) per 100,000 eligible voters.

onesided -> One-sided deaths (civdeath_gov + civdeath_mao) per 100,000 eligible voters.

## Plots

We can visualize how the Maoist vote share moves with gender and the age of the winning candidate in each electoral district.

```{r}
ggplot(df, aes(winnerage, maovote, colour = winnergender)) + 
  geom_point() +  
  geom_smooth(method=lm, se=FALSE) + labs(x = "Age of Winning Candidate", y = "Maoist Vote", color = "Gender")
```

This plot visualizes the relationship between the average age of the winning candidate and voter turnout separated by gender. We can see that there is a much stronger positive relationship between winner age and turnout in constituencies with female winning candidates as a pose to electoral districts with male winners.

```{r}
ggplot(df, aes(winnerage, turnout, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE) + labs(x = "Age of Winning Candidate", y = "Turnout", color = "Gender")
```

This plot depicts the association between the average age of the candidates contesting the single-member constituencies in the 2008 elections and the Maoist vote share. On average, as the average age of the candidates increased,the Maoist vote share in the electoral districts increased at a roughly similar rate in constituencies won by men and women.

```{r}
ggplot(df, aes(candidate_avgage, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE) + labs(x = "Average Age of Candidate", y = "Maoist Vote", color = "Gender")
```


## Gender in the Nepalese Civil War

A topic I did not adequately touch on in my thesis was the the prominent role played by women cadre and fighters in the Maoist movement. Based on the data, one-eighth of the winning candidates in the electoral districts were women.  

```{r}
ggplot(df, aes(winnergender, fill=winnergender)) + 
  geom_bar(width = 0.5) + theme_minimal() + labs(x = "Gender of Winning Candidates", y = "Number of Winning Candidates", fill = "Gender")

```


Let's examine some trends related to violence and Maoist electoral performance through the lens of the gender of the winning candidates. An observation that holds across all of the predictor variables was the inherently stronger support for the CPN(M) in the constituencies won by women. When observing the effect of civilian atrocities perpetrated by government forces on support for the Maoists in the 2008 elections, something distinct stands out. On average, it appears that the electoral districts with female winning candidates had less violence to begin with, but voted for the Maoist candidate in greater numbers. However, in spite of the small sample size, we can observe that there exists a much more positive association between government violence and Maoist electoral success in constituencies won by women compared to those won by men.

```{r}
ggplot(df, aes(civdeath_gov, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE) + labs(x = "Civilian Deaths Perpetrated by Government Forces", y = "Maoist Vote", color = "Gender")
```

It is also prudent to observe the relationship between turnout and Maoist support. Election observers like the Carter Center noted that the Maoists engaged in the intimidation and harassment of their opponents in the days leading up to the polls, but that the election itself took place in reasonably tranquil conditions. Here, we can see that regardless of gender, the Maoists performed better in electoral districts where fewer eligible voters turned out. There appears to be a stronger negative relationship between turnout and Maoist electoral performance for constituencies where women were victorious versus those won by male candidates.

```{r}
ggplot(df, aes(turnout, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE) + labs(x = "Turnout", y = "Maoist Vote", color = "Gender")
```

## Political Parties in the 2008 Constituent Assembly Election

```{r}
ggplot(df, aes(x = winningparty, fill=winningparty)) + 
  geom_bar(width = 0.5) + theme_minimal() +
  scale_fill_manual(values = c("darkred", "grey", "orange", "orangered", "darkgreen" , "#ff8080" , "seagreen" , "firebrick" , "lightgrey" , "red"))+ 
  theme(axis.text=element_text(size=10)) + labs(x = "Winning Party", y = "Number of Seats", fill = "Party")
```


Let's calculate the effective number of parties per electoral district. 


```{r}

library(knitr)
library(dplyr)

df2 <- read.csv("partylistresults.csv")

df2 <- as.data.frame(df2)

p <- df2[, -(1:4)]

calculate_N <- function(row) {
  sum_squares <- sum(row^2)
  N <- 1 / sum_squares
  return(N)
}

N_values <- apply(p, 1, calculate_N)

df2 <- cbind(df2[, 4, drop = FALSE], N = round(N_values, 2))

df2 <- df2 %>%
  mutate(across(where(is.numeric), ~ round(., 2)))

kable(head(df2))


```



Here is a chart that regresses turnout in the constituent assembly election and the effective number of parties per electoral district, disaggregated by gender.

```{r}
df3 <- read.csv("partylistresults.csv")
df3$N <- N_values
df4 <- read.csv("completedata.csv")

ggplot(df3, aes(df4$turnout, N, colour = df4$winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE) +
  theme(axis.text = element_text(size = 6)) + labs(title = "Effective Number of Parties and Turnout" , x = "Turnout" , y = "Effective Number of Parties", color = "Gender")
```
