---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Nepal-Conflict-and-Elections

<!-- badges: start -->

<!-- badges: end -->


## Introduction

Nepal is one of only a few countries in the world that successfully integrated a rebel group into its political system, a process that has yielded relatively peaceful results to the present day. From 1996 to 2006, the Nepalese people faced the tragedy of a [civil war](https://en.wikipedia.org/wiki/Nepalese_Civil_War) that destroyed lives, communities and families alike. For the personal stories of those affected, the [INSEC](https://www.insec.org.np/victim) database comprehensively catalogues details on the thousands of dead, wounded and missing of the conflict.



Given Nepal's unique story, I put together a study to examine the relationship between a rebel group's actions during a civil conflict (regarding their treatment of civilians and casualties inflicted on government forces), and its performance in postwar transitional elections. 


## The data

Here is a subsection of the dataset I compiled. There is also data on Maoist electoral performance, and some attributes of the successful and unsuccessful candidates who contested the single-member constituencies. 

```{r}
library(tidyverse)
library(kableExtra)

df = read.csv("completedata.csv")
kable(head(df))
```



## Descriptive statistics


With this data, we can examine the mean age of candidates and winning candidates in each electoral district along with concurrent violence statistics.

```{r}
mean(df$winnerage)
mean(df$candidate_avgage)
mean(df$civdeath_gov)
mean(df$civdeath_mao)
mean(df$govtdeath)
mean(df$maodeath)
mean(df$statebased)
mean(df$onesided)




```
Key

civdeath_gov -> Civilian deaths perpetrated by Government Forces (Royal Nepalese Army, Nepal Police, Armed Police Force) per 100,000 eligible voters

civdeath_mao -> Civilian deaths perpetrated by the Communist Party of Nepal (Maoist) per 100,000 eligible voters

govtdeath -> Government deaths perpetrated by the Maoist rebels per 100,000 eligible voters

maodeath -> Maoist deaths perpetrated by Government forces per 100,000 eligible voters.

statebased -> State-based deaths (govtdeath + maodeath) per 100,000 eligible voters.

onesided -> One-sided deaths (civdeath_gov + civdeath_mao) per 100,000 eligible voters.

## Plots

We can visualize how the age of winning candidates moves with gender and the Maoist vote share in each electoral district.

```{r}
ggplot(df, aes(maovote, winnerage, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```


This plot regresses turnout with the average age of the winning candidate separated by gender. We can see that there is a much stronger positive relationship between turnout and winner age in constituencies with female winning candidates as a pose to male candidates.

```{r}
ggplot(df, aes(turnout, winnerage, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```


This plot regresses Maoist performance with the average age of the candidates contesting single-member constituencies disaggregated by gender.

```{r}
ggplot(df, aes(maovote, candidate_avgage, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

Now, moving onto regressions that involve conflict statistics. My study's main hypothesis argued that there should be a negative association between the level of atrocities the Maoists committed and their electoral performance, and a positive association between the casualty rate inflicted on government forces and Maoist electoral performance. 


## Gender in the Nepalese Civil War

A topic I did not adequately touch on in my thesis was the the prominent role played by women cadre and fighters in the Maoist movement. Let's examine some trends of violence and Maoist electoral performance through the lens of the gender of the winning candidates. A general trend across all of the predictor variables was the inherently stronger support for the CPN(M) in the constituencies won by women. 

In this first chart, we can see that the few constituencies with female winning candidates generally had greater Maoist support, but that there is a more positive relationship between Maoist atrocities and Maoist electoral performance in electoral districts where the winning candidates were women.

```{r}
ggplot(df, aes(civdeath_mao, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

By contrast, when studying the effect of civilian deaths caused by government forces on Maoist electoral performance, the differences by gender of the winning candidate are far more drastic.

```{r}
ggplot(df, aes(civdeath_gov, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

We can see that the Maoist vote share increased markedly along with the death rate of government forces. In addition to the generally higher Maoist vote shares in constituencies won by women, it is again noticeable that the rebel vote share increased by more in these electoral districts relative to the male ones. 


```{r}
ggplot(df, aes(govtdeath, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

Here is the relationship between rebel deaths and rebel vote share in the 2008 elections. In this case, the slope of the regression line for constituencies with male winners is steeper than the regression for the few constituencies where women were elected

```{r}
ggplot(df, aes(maodeath, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```

Finally, an examination of turnout and Maoist support. It is noted by election observers like the Carter Center, that the Maoists engaged in intimidation and harassment in the leadup to the polls, but that the election itself took place in reasonably tranquil conditions. Here, we can see that regardless of gender, the Maoists performed better in electoral districts where fewer eligible voters turned out. Once again, the general effect is even more pronounced in constituencies where women were victorious.

```{r}
ggplot(df, aes(turnout, maovote, colour = winnergender)) + 
  geom_point() + 
  geom_smooth(method=lm, se=FALSE)
```
